{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-19 thesis/future.tex
tweak wording of dunk on machine learning in future

%% commenting this first part out bc it will probably aggro garth...
%As a matter of personal philosophy,
%I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set
(or in the minds of its developers)
in unpredictable ways
that can harm people of color \cite{conceptnet-bias,gender-shades},
women \cite{gender-bias-ai,amazon-ai},
and transgender people \cite{misgendering-machines}.
In the domain of concurrency testing,
%careless application of
such bias would be unlikely to result in interpersonal discrimination,
but
%I believe
it would translate to
%being able to find
finding only types of bugs
already seen in the training data,
abandoning any novel bug patterns to further obscurity.
Granted,
Iterative Deepening and
ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
%in terms of minimum necessary preemption count,
%but this strategy is informed by empirical evidence
%% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
%\cite{chess-icb},
%and
but when a bug is not found
these strategies also enable the user to understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\cref{sec:tm-eval})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
I firmly believe that using machine learning as a bug-finding heuristic
without compensating extremely carefully for its inherent biases
would run counter to this thesis's central tenets of transparency and user agency.
Any overall search heuristic
%for concurrency testing
must be designed to support the human user first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.

>>>>

%% commenting this first part out bc it will probably aggro garth...
%As a matter of personal philosophy,
%I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set
(or in the minds of its developers)
in unpredictable ways
that have been shown to harm people of color \cite{conceptnet-bias,gender-shades},
women \cite{gender-bias-ai,amazon-ai},
and transgender people \cite{misgendering-machines}.
In the domain of concurrency testing,
%careless application of
such bias would be unlikely to result in interpersonal discrimination,
but
%I believe
it would translate to
%being able to find
finding only types of bugs
already seen in the training data,
abandoning any novel bug patterns to further obscurity.
Granted,
Iterative Deepening and
ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
%in terms of minimum necessary preemption count,
%but this strategy is informed by empirical evidence
%% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
%\cite{chess-icb},
%and
but when a bug is not found
these strategies also enable the user to understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\cref{sec:tm-eval})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
I firmly believe that using machine learning as a bug-finding heuristic
without compensating extremely carefully for its inherent biases
would run counter to this thesis's central tenets of transparency and user agency.
Any overall search heuristic
%for concurrency testing
must be designed to support the human user first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.
